---
title: rCharts
subtitle: Interactive Charts from R using Polycharts
author: Ramnath Vaidyanathan
github: {user: ramnathv, repo: rCharts, branch: "gh-pages"}
framework: minimal
mode: selfcontained
widgets: polycharts
hitheme: solarized_light
---

## Interactive Charts from R using [PolyCharts](https://github.com/Polychart/polychart2)

This R package helps users create interactive visualizations using [polychart.js](https://github.com/Polychart/polychart2), using a plotting interface similar to the [lattice](http://cran.r-project.org/web/packages/lattice/index.html) package.

### Installation

You can install `rCharts` from `github` using the `devtools` package

```
require(devtools)
install_github('rCharts', 'ramnathv')
```

### Usage

`rCharts` uses a plotting interface similar to that of the `lattice` package. Here are a few examples you can try out in your R console.

### Example 1: Faceted Scatterplot

<div id='chart1' class='rChart'></div>

```{r results = 'asis', comment = NA, message = F}
require(rCharts)
names(iris) = gsub("\\.", "", names(iris))
p1 <- rPlot(SepalLength ~ SepalWidth | Species, data = iris, color = 'Species', 
  type = 'point')
p1$addParams(width = 550)
p1$printChart('chart1') # use p1$show() from your R console 
```

### Example 2: Faceted Barplot

<div id='chart2' class='rChart'></div>

```{r results = 'asis', comment = NA, message = F, tidy = F}
hair_eye = subset(as.data.frame(HairEyeColor), Sex == "Female")
p2 <- rPlot(Freq ~ Hair, color = 'Eye', data = hair_eye, type = 'bar', 
 position = 'dodge')
p2$addParams(width = 550)
p2$printChart('chart2')  # use p2$show() from your R console
```

### Example 3: Faceted Grid

<div id='chart3' class='rChart'></div>

```{r results = 'asis', comment = NA, message = F, tidy = F}
p3 <- rPlot(mpg ~ wt | vs + am, data = mtcars, type = 'point', color = 'gear')
p3$addParams(width = 550)
p3$printChart('chart3')
# qplot(wt, mpg, data = mtcars, facets = vs ~ am, color = gear, geom = 'point')
```


`rCharts` is merely a wrapper around Polycharts.js. More documentation and examples are underway.

### Using with Shiny

rCharts is compatible with Shiny. Here is a link to the [rCharts Shiny App](http://glimmer.rstudio.com/ramnathv/rChartApp/).

```r
## server.r
require(rCharts)
shinyServer(function(input, output) {
  output$myChart <- renderChart({
    names(iris) = gsub("\\.", "", names(iris))
    p1 <- rPlot(input$x, input$y, data = iris, color = "Species", 
      facet = "Species", type = 'point')
    p1$addParams(dom = 'myChart')
    return(p1)
  })
})

## ui.R
require(rCharts)
shinyUI(pageWithSidebar(
  headerPanel("rCharts: Interactive Charts from R using polychart.js"),
  
  sidebarPanel(
    selectInput(inputId = "x",
     label = "Choose X",
     choices = c('SepalLength', 'SepalWidth', 'PetalLength', 'PetalWidth'),
     selected = "SepalLength"),
    selectInput(inputId = "y",
      label = "Choose Y",
      choices = c('SepalLength', 'SepalWidth', 'PetalLength', 'PetalWidth'),
      selected = "SepalWidth")
  ),
  mainPanel(
    showOutput("myChart", "polycharts")
  )
))
```

### Credits

Most of the implementation in `rCharts` is inspired by [rHighcharts](https://github.com/metagraf/rHighcharts) and [rVega](https://github.com/metagraf/rVega). I have reused some code from these packages verbatim, and would like to acknowledge the efforts of its author [Thomas Reinholdsson](https://github.com/reinholdsson).

### License

`rCharts` is licensed under the MIT License. However, the Polycharts JavaScript library that is included in this package is not free for commercial use, and is licensed under Creative Commons 3.0 Attribution & Non-commercial. Read more about its license at http://polychart.com/js/license.



