---
title: rCharts
subtitle: Interactive Charts from R using Polycharts
author: Ramnath Vaidyanathan
github:
  user: ramnathv
  repo: rCharts
framework: minimal
mode: selfcontained
widgets: polycharts
hitheme: solarized_light
---

## Interactive Charts from R using [PolyCharts](https://github.com/Polychart/polychart2)

This R package helps users create interactive visualizations using [polychart.js](https://github.com/Polychart/polychart2), using a plotting interface similar to the [lattice](http://cran.r-project.org/web/packages/lattice/index.html) package.

### Installation

You can install `rCharts` from `github` using the `devtools` package

```
require(devtools)
install_github('rCharts', 'ramnathv')
```

### Usage

`rCharts` uses a plotting interface similar to that of the `lattice` package. Here are a few examples you can try out in your R console.

### Example 1: Faceted Scatterplot

<div id='chart1' class='rChart'></div>

```{r results = 'asis', comment = NA, message = F}
require(rCharts)
names(iris) = gsub("\\.", "", names(iris))
p1 <- rPlot(SepalLength ~ SepalWidth | Species, data = iris, color = 'Species', 
  type = 'point')
p1$addParams(width = 550)
p1$printChart('chart1')
```

### Example 2: Faceted Barplot

<div id='chart2' class='rChart'></div>

```{r results = 'asis', comment = NA, message = F, tidy = F}
hair_eye = subset(as.data.frame(HairEyeColor), Sex == "Female")
p2 <- rPlot(Freq ~ Hair, color = 'Eye', data = hair_eye, type = 'bar', 
 position = 'dodge')
p2$addParams(width = 550)
p2$printChart('chart2')
```



