## R Notebook and rCharts

This is a short demo on how to embed d3js style charts in an R Markdown notebook, using rCharts and Shiny. If you want to follow along, please install `rCharts` from `github` using `devtools` and run `open_notebook()`.

```{r eval = F}
require(devtools)
install_github('rCharts', 'ramnathv')
rCharts::open_notebook()
```

Before we begin charting, let us customize some settings for rCharts and knitr

```{r echo = T, message = F, cache = F}
require(rCharts)
options(RCHART_WIDTH = 600, RCHART_HEIGHT = 400)
knitr::opts_chunk$set(comment = NA, results = 'asis', tidy = F, message = F)
```

<style>.rChart {width: 600px; height: 400px;}</style>

Time to get charting now!

---

## Polychart

We will create our first chart using Polychart, a javascript charting library based on the grammar of graphics, and inspired by ggplot2.

```{r chart1}
r1 <- rPlot(mpg ~ wt | am + vs, data = mtcars, type = 'point',
  color = 'gear')
r1$print('chart1')
```

There, we have our first embedded chart with nice tooltips! Let me add some interactivity to this chart now using javascript.

<script>
graph_chart1.addHandler(function(type, e){
  var data = e.evtData;
  if (type === 'click'){
    return alert("You clicked on car with mpg: " + data.mpg.in[0]);
  }
})
</script>

---

## Morris

The next library we will be exploring is MorrisJS.

```{r chart2}
data(economics, package = 'ggplot2')
econ <- transform(economics, date = as.character(date))
m1 <- mPlot(x = 'date', y = c('psavert', 'uempmed'), type = 'Line',
  data = econ)
m1$set(pointSize = 0, lineWidth = 1)
m1$print('chart2')
```

Hurray! There we have our second chart!

---

## NVD3

Next, I will demonstrate my all time favorite d3js library, NVD3, which produces amazing interactive visualizations with little customization.

```{r chart3}
hair_eye_male <- subset(as.data.frame(HairEyeColor), Sex == "Male")
n1 <- nPlot(Freq ~ Hair, group = "Eye", data = hair_eye_male, 
  type = 'multiBarChart')
n1$print('chart3')
```

See the interactivity that comes at zero cost! 

---

## xCharts

Finally, I will demo xCharts, a slick looking charting library using d3js, made by TenXer.

```{r chart4}
require(reshape2)
uspexp <- melt(USPersonalExpenditure)
names(uspexp)[1:2] = c('category', 'year')
x1 <- xPlot(value ~ year, group = 'category', data = uspexp, 
  type = 'line-dotted')
x1$print('chart4')
```

There is your xChart

---

## HighCharts

```{r chart5}
h1 <- hPlot(x = "Wr.Hnd", y = "NW.Hnd", data = MASS::survey, 
  type = c("line", "bubble", "scatter"), group = "Clap", size = "Age")
h1$print('chart5')
```

---

## Leaflet

```{r}
map3 <- Leaflet$new()
map3$setView(c(51.505, -0.09), zoom = 13)
map3$marker(c(51.5, -0.09), bindPopup = "<p> Hi. I am a popup </p>")
map3$marker(c(51.495, -0.083), bindPopup = "<p> Hi. I am another popup </p>")
map3$print('chart7')
```

---

## Rickshaw

```{r chart6}
usp = reshape2::melt(USPersonalExpenditure)
p4 <- Rickshaw$new()
p4$layer(value ~ Var2, group = 'Var1', data = usp, type = 'area')
p4$print('chart6')
```

---

## Credits

I would like to acknowledge the efforts of @yihui and @rstudio in bringing R Notebooks to the R community. I would like to thank @yihui for his excellent shiny app in the knitr package, which was the basis for this app. Without their efforts, rCharts and this demo would not have been possible.

For more information, please visit

http://github.com/ramnathv/rCharts!
